---
title: "Join and export data"
output: html_notebook
---

In this script I join and export the raw data from the experiments. 

We have data from 4 different antibiotics + control conditions (no antibiotics) in triplicates. 

We have the following antibiotics:

+ Chloramphenicol (Cam)
+ Kanamycin A (Kan)
+ Tetracycline (Tet)
+ Trimethoprim (Trim)

Each plate comes with a separate result file, store in Batch folders by experiment. 
The Data for Kan have been repeated and are joined from another batch.

```{r, message=FALSE, warning=FALSE}

path <- here("Data", 
             "Results_Exp3",
             "190904_mix5_QC_normalised_absolute_batch")

Files <- list.files(path, pattern = ".csv")

#exclude Kan from experiment 3
Files <- Files[!grepl("kan8", Files)]

File_list <- 
lapply(Files, function(x) {
  temp <- read_csv(paste(path, x, sep = "/"))
  name <- sub("^.+_(\\w+\\d)_.+","\\1", x)
  Name <- paste(toupper(substr(name,1,1)), substr(name,2,nchar(name)), sep ="")
  tempm <- mutate(temp,
                  Plate = Name,
                  Antibiotic = gsub("([a-zA-Z]+).+", "\\1", Name),
                  Replicate = substr(Name, nchar(Name), nchar(Name)))
  })

# load Kan from experiment 4

path2 <- here("Data", 
             "Results_Exp4_Kan",
             "Kan6",
             "191108_Lars_Kan6_Kan6_Kan6_M9_Jaime.phenotypes.NormalizedAbsoluteBatched")

Files2 <- list.files(path2, pattern = ".csv")

#exclude additional controle (Plate 4)
Files2 <- Files2[!grepl("plate_4", Files2)]

File_list2 <- 
lapply(Files2, function(x) {
  temp <- read_csv(paste(path2, x, sep = "/"))
  Name <- paste("Kan6", gsub(".+(\\d).csv", "\\1", x), sep = "_")
  tempm <- mutate(temp,
                  Plate = Name,
                  Antibiotic = gsub("([a-zA-Z]+).+", "\\1", Name),
                  Replicate = substr(Name, nchar(Name), nchar(Name)))
  })

```

## clean data

+ bind all dataframes together to one long dataframe
+ export joined data and uplaod to figshare. 


```{r}

Data <-  do.call("rbind", c(File_list, File_list2))

#export Data to upload to Figshare
write_tsv(Data, here("Data", "Gamfeldt_et_al_Biodiversity_and_EF_across_scale.txt"))
```


These raw data have been reposited to fishare and are citable with the following reference:

> Gamfeldt, Lars; Roger, Fabian; Palm, Martin; Hagan, James; Warringer, Jonas; Farewell, Anne (2020): Biodiversity and ecosystem functioning across gradients in spatial scale. figshare. Dataset. https://doi.org/10.6084/m9.figshare.12279884.v1


You can download the data from here: 
[https://doi.org/10.6084/m9.figshare.12279884.v1](https://doi.org/10.6084/m9.figshare.12279884.v1)
