---
title: "script to make multipanel figure 2 and S3"
output: html_notebook
---

```{r}
library(ggplot2)
library(here)
library(cowplot)
source(here("function_plotting_theme.R"))
```



#Figure 2

To produce this figure, you must run the code for Figure 2 a,b & Figure 2 a,b insets in Siumlations.Rmd and the code for Figure 2 c,d in Analysis.nb.Rmd. The `Scenario` in Simulations must be set to 2 (weighted mean) and the sensitivity analysis  (Analysis) must be toggled of. 

load figures
```{r}

# load(here("figures", "Fig_2_a.RData"))
load(here("figures", "Fig_2_b.RData"))
# load(here("figures", "Fig_2_a_i.RData"))
load(here("figures", "Fig_2_b_i.RData"))

# load(here("figures", "Fig_2_c.RData"))
load(here("figures", "Fig_2_d.RData"))
```


arrange figures
```{r}
# Fig_2_a <- 
# Fig_2_a +
  # scale_y_continuous(limits = c(-0.15, NA))+
  # theme_meta()+
  # labs(y = "BEF slope", x = "")


Fig_2_b <- 
  Fig_2_b +
  labs(x = "")+
  theme_meta()+
  scale_y_continuous(limits = c(-0.15, NA))

# Fig_2_c <- Fig_2_c +
  # theme_meta()+
  # labs(y = "BEF slope")


Fig_2_d <- 
  Fig_2_d+
  labs(x = "")+
   theme_meta()

# Fig_2_a_i <- Fig_2_a_i+
  # labs(x = "", y = "", title = "")+
  # theme_minimal()+
  # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        # axis.text=element_text(size=8, face = "bold"),
        # axis.title=element_text(size=8, face = "bold"),
        # panel.border = element_rect(colour = NA, fill=NA),
        # axis.ticks.y = element_blank(),
        # axis.text.y = element_blank())+
  # scale_x_continuous(limits = c(NA, NA))+
  # theme(axis.text.x = element_text(angle = -45, hjust = 0, vjust = -0.05))


Fig_2_b_i <- 
  Fig_2_b_i+
  labs(x = "", y = "", title = "") #+
  # theme_minimal()+ 
  # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        # axis.text=element_text(size=8, face = "bold"),
        # axis.title=element_text(size=8, face = "bold"),
        # panel.border = element_rect(colour = NA, fill=NA),
        # axis.ticks.y = element_blank(),
        # axis.text.y = element_blank())+
  # theme(axis.text.x = element_text(angle = -45, hjust = 0, vjust = -0.05))


# Fig_2_a <- Fig_2_a +
  # annotation_custom(
    # ggplotGrob(Fig_2_a_i),
   # xmin = 2, xmax = 5, ymin = -0.2, ymax = 0.1
  # )

Fig_2_b <- 
  Fig_2_b +
  annotation_custom(
    ggplotGrob(Fig_2_b_i),
    xmin = 2, xmax = 5, ymin = -0.3, ymax = 0.6
  )


# p1 <- plot_grid(Fig_2_a, Fig_2_b,
                # Fig_2_c, Fig_2_d, 
                # align = "v",
                # labels="AUTO",
                # label_x = 0.02)

# ggsave(plot = p1, filename = here("figures", "Fig_2.pdf"), 
       # width = 15.24, height = 17, units = "cm", dpi = 450)

# add another figure without the BEF slope
p2 <- ggpubr::ggarrange(Fig_2_b + labs(x = "Scale", y = "Transgressive overyielding"),
                        Fig_2_d + labs(x = "Scale", y = "Transgressive overyielding"), 
                        ncol = 2, nrow = 1, widths = c(1, 1),
                  labels = c("A", "B"),
                  font.label = list(size = 10, color = "black", face = "bold", family = NULL))

p2

ggsave(filename = here("figures/fig_2.pdf"), 
       plot = p2, width = 18, height = 9, units = "cm", dpi = 450)

```

#Figure 3

To produce this figure, you must run the code for Figure 3 a,b & in Analysis.nb.Rmd. The sensitivity analysis  (Analysis) must be toggled of. 

load Figures
```{r}
load(here("figures", "Fig_3A.RData"))
load(here("figures", "Fig_3B.RData"))
```


arrange Figures
```{r}
Fig_3B <- 
  Fig_3B+
  theme_meta()+
  xlab("Strain richness") +
  theme(legend.position = "none")+
  labs(y = "")

Legend <- get_legend(
  # create some space to the left of the legend
  Fig_3A + 
    theme(legend.position = "bottom", 
          legend.box.margin = margin(0, 0, 0, 12),
          legend.key=element_blank()) 
)

Fig_3A <- 
  Fig_3A+
  theme_meta()+
  theme(legend.position = "none")

# add the legend to the row we made earlier. Give it one-third of 
# the width of one plot (via rel_widths).

Fig3 <- plot_grid(Fig_3A, Fig_3B,
                align = "v",
                labels="AUTO",
                label_x = 0.1, 
                label_size = 10)

Fig3 <- plot_grid(Fig3, Legend, nrow = 2, rel_heights = c(2, .4))

Fig3

ggsave(plot = Fig3, filename = here("figures", "Fig_3.pdf"), 
       width = 18, height = 9, units = "cm", dpi = 450)
```


#Figure S8

note that to produce this figure you must run the code in Analysis.nb.Rmd both with and without the sensitivity analysis

load figures
```{r}
load(here("figures", "Fig_2_d.RData"))

Fig_2_d_with_a <- 
  Fig_2_d+
  theme_meta()+
  scale_y_continuous(limits = c(0,1.1))

load(here("figures", "Fig_S_9.RData"))

Fig_2_d <- Fig_2_d+
   theme_meta()+
  scale_y_continuous(limits = c(0,1.1))
```


arrange figures
```{r}
p1 <- plot_grid(Fig_2_d_with_a, Fig_2_d, 
                align = "v", labels="AUTO",
                label_x = 0)

p1

ggsave(plot = p1, filename = here("figures", "Fig_S8.pdf"), width = 8, height = 4)


```
