---
title: "Bacterial EF across Scale - Simulations"
output: html_notebook
---

```{r, message=FALSE}
library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)
library(ggbeeswarm)
library(broom)
library(viridis)
library(here)
library(pbapply)
library(nlstools)
```

This script exports: 

+ the subplots for Figure 2 (Fig 2 a and b as well as the insets)
+ Fig S8

See script `Multipanel_Fig_2.Rmd` for the assemblage and export of figure 2. 

make folder to export figures
```{r}

if(! dir.exists(here("figures"))){
  dir.create(here("figures"))
}

```


#set up simulations

+ We want to simulate 5 strains of Bacteria, growing in 5 different environments. 
+ Each Strain differs in its ability to grow on each antibiotic.
+ We grow the bacteria in every possible combination (monocultures and all possible species mixtures)
+ We grow them on every possible combination of environments
+ We measure EF at different scales going from 1 to 5 environments  

define number of species and environments
```{r}

specnum <- 5
envnum <- 5
 
```

## strains

We create 5 species (for `specnum = 5`) named strain_1 to strain_5 and get every possible combination of the five species for all 1,2,3,4 and 5 species combination (corresponding richness levels 1:5 in a biodiversity experiment)

```{r}
spec <- paste("strain_", c(1:specnum), sep = "")
spec_comb <- sapply(c(1:specnum), function(x) combn(spec, x)) 
names(spec_comb) <- paste("richness", c(1:specnum))
```

## environments

We create 5 environments (for `envnum = 5`) named Environment_1 to Environment_5 and get every possible combination of the five environments for all 1,2,3,4 and 5 environment combinations (corresponding to the different scales)

```{r}
env <- paste("Environment_", c(1:envnum), sep = "")
env_comb <- sapply(c(1:envnum), function(x) combn(env, x)) 
names(env_comb) <- paste("Heterogeneity", c(1:envnum))
```

number of combinations

```{r}
tibble(spec_combs = 
         sum(choose(specnum,1:specnum)),
       env_combs = 
         sum(choose(envnum,1:envnum)))
```

# step-by-step

In the following we draw a single draw of fitness values for each strain and develop the code step by step. The code that generates the results and figures presented in the paper relies on 1000 different draws and is found under the section `multiple draws from uniform`

### fitness values

We assign each strain a fitness value for each environment. The values are drawn from a uniform distribution (0,1) for each strain and each environment. 

```{r}

set.seed(1346)

Traits <- matrix(
  nrow = length(env),
  ncol = length(spec),
  dimnames = list(env, spec))


Traits[1:nrow(Traits), 1:ncol(Traits)] <-
  runif( n = prod( dim( Traits)), 0, 1)

Traits
```

### calculate plot values

here we calculate the outcome (biomass / 'functioning') of every species combination in each environment. We calculate the mixture performance as:

**Weighted mean**: The mixture value equals the mean performance of all species, weighted by their relative performance in the environment

*what the code does*

for each environment `sapply(env, ...` it cycles through all richness levels `lapply(spec_comb, ...)` and for each richness level it cycles through all species combinations `apply(..., 2, )`.
Then it calculates the mixture performance of that species combination of that richness level in that environment given the species performance stored in `Traits` 

```{r}

plot_values <- sapply(env, #for each environment
                      function(x) {
                        lapply(spec_comb, #for each richness level
                               function(y) {
                                 apply(y, 2, function(z){ #for each species combination
                                   mean(sum(Traits[x,z]^2) / abs(sum(Traits[x,z]))) #weighted mean
                                   }
                                 )
                                 }
                        )
                        }
                      )

plot_values_df <- 
  tibble(
    richness = rep( rep(1:specnum, unlist(lapply(spec_comb, ncol))), envnum),
    spec_comb = rep(lapply(spec_comb, 
                           function(x) apply(x, 2, function(y) 
                             paste(y, collapse = " "))) %>% unlist(),
                    envnum),
    environment = rep(env, each = sum(choose(specnum,1:specnum))),
    functioning = unlist(plot_values)
    )

#small random error so that the function fitting functions don't fail

# plot_values_df <- 
#   plot_values_df %>% 
#   mutate(functioning = functioning + rnorm(n(), 0, 0.01)) %>% 
#   mutate(functioning = functioning + abs(min(functioning)) + 0.01)

```

### plot plot-values

Here we plot the single species and mixture functioning of each species combination in each environment. The Mixture performance is calculated as the weighted mean of the monoculture performances. 

We show a power-fit (with 2*se CI) and a linear fit, both showing the positive richness ~ functioning relationship (that follows inherently from the way we calculate the mixture values)

```{r}

plot_values_df %>% 
  na.omit() %>% 
  group_by(environment) %>%
  nest() %>% 
  mutate(power_fit = map(data, ~lm(log(functioning) ~ log(richness), data = .))) %>% 
  mutate(power_fit = map(power_fit, augment, se_fit = TRUE)) %>% 
  unnest(cols = c(data, power_fit)) %>% 
  ggplot(aes(x = richness, y = functioning))+
  geom_point(position = position_jitter(width = 0.1, height = 0), 
             alpha = 1, size = 1, colour = "#377EB8")+
  geom_line(aes(x = richness, y = exp(.fitted)), 
            colour = "black", size = 0.7)+
  geom_smooth(method = "lm", se = F, linetype = "dotted", size = 0.7, colour = "darkred")+
  geom_ribbon(aes(x = richness, 
                  ymin = exp(.fitted - 2*.se.fit), 
                  ymax = exp(.fitted + 2*.se.fit),
                  group = environment), fill = "grey", alpha = 0.2)+
  facet_wrap(~environment, scales = "free_y")+
  theme_bw()

```

## BEF metrics

We calculate the slope between richness and functioning for all combinations of environments. 

+ for all scales (1:5 environments)
+ for all possible unique combinations of environments 

We calculate the functioning of a given species mixture in that combination of environments at that scale as the average functioning of that species mixture in the single environments. 

For each scale and environment combination we then calculate the following metrics:

+ the **slope of a power fit** to the richness ~ functioning relationship
+ **transgressive overyielding**: highest mixture value / maximum monoculture value


```{r, warning=FALSE}

plot_values_wide <- 
plot_values_df %>% 
  spread(environment, functioning)

#calculate average functioning for each env combination at each scale and store results in dataframe
env_values <- 
lapply(env_comb, function(x) { #for all scales (1:5)
  apply(x, 2, function(y) { #for all environmental combinations
   
     df <- data.frame( #average functioning for this environment combination
       env_comb = paste0(y, collapse = " "),
       envnum = length(y),
       richness = plot_values_wide$richness,
       functioning = rowMeans(plot_values_wide[,y]) #mean of functioning across environments
                     )
  })
  })

env_values <- bind_rows(unlist(env_values, recursive = FALSE))

# calculate BEF metrics for each environment combination
env_metrics <- 
  env_values %>% 
  group_by(env_comb, envnum) %>%
  nest() %>% 
  mutate(power_fit = map(data, ~nls(functioning ~ a*richness^b, data = ., start = list(a = 1, b = 1)))) %>%
  mutate(power_fit = map_dbl(power_fit, function(x) coef(x)[2])) %>%
  mutate(transgressive = map_dbl(data, function(x) {
    x[x$richness == specnum,]$functioning / 
      max(x[x$richness == 1,]$functioning)})) 
```

### plot power-fits 

```{r}

env_values %>%
  group_by(env_comb, envnum) %>% 
  nest() %>% 
  mutate(power_fit = map(data, ~nls(functioning ~ a*richness^b, data = ., start = list(a = 1, b = 1)))) %>%
  #mutate(power_fit = map(data, ~nls(functioning ~ a*richness/(b+richness), data = ., start = list(a = 1, b = 1)))) %>%
  mutate(power_fit = map(power_fit, augment)) %>% 
  unnest(cols = power_fit) %>% 
  ggplot(aes(x = richness, y = functioning, colour = as.factor(envnum)))+
  geom_point(aes(group = env_comb), 
             size = 1, alpha = 1, 
              position = position_dodge(width = 0.7)
              )+
  geom_line(aes(y = .fitted, group = env_comb),
            position = position_dodge(width = 0.7)
            )+
  facet_wrap(~envnum)+
  scale_color_manual(values = viridis(envnum+1)[1:envnum], name = "scale")+
  theme_classic()+
  labs(y = "production", x = "species richness")+
  theme(legend.position = "none")
  

```



### plot power-fits by scale

```{r}

env_metrics %>% 
  select(-data) %>% 
  gather(metric, value, -envnum, -env_comb) %>%
  filter(metric == "power_fit") %>% 
  #filter(metric == "MichMent") %>% 
  ggplot(aes(x = envnum, y = value))+#, colour = as.factor(envnum)))+
  geom_point( size = 2, alpha = 1, position = position_jitter(width = 0.1), colour = viridis(6)[3])+
  stat_smooth(method = "lm", se = F, size = 0.5, colour = "darkred", linetype = "dashed")+
 # scale_color_manual(values = viridis(envnum+1)[1:envnum], name = "scale")+
  theme_classic()+
  labs(y = "beta_slopes", x = "scale")+
  theme(legend.position = "none",
        axis.ticks.length=unit(-0.25, "cm"),
        axis.text.x = element_text(margin=unit(c(0.5,0.5,0.5,0.5), "cm")),
        axis.text.y = element_text(margin=unit(c(0.5,0.5,0.5,0.5), "cm")) )+
  scale_y_continuous(limits = c(0,0.26))+
  NULL


```


### plot transgressive overyielding
```{r}

env_metrics %>% 
  select(-data) %>% 
  gather(metric, value, -envnum, -env_comb) %>% 
  filter(metric == "transgressive") %>% 
  ggplot(aes(x = envnum, y = value))+
  geom_point( size = 2, alpha = 1, position = position_jitter(width = 0.1), colour = viridis(6)[3])+
  stat_smooth(method = "lm", se = F, size = 0.5, colour = "darkred")+
  theme_classic()+
  labs(y = "transgressive overyielding", x = "scale")+
  theme(legend.position = "none",
        axis.ticks.length=unit(-0.25, "cm"),
        axis.text.x = element_text(margin=unit(c(0.5,0.5,0.5,0.5), "cm")),
        axis.text.y = element_text(margin=unit(c(0.5,0.5,0.5,0.5), "cm")) )+
  scale_y_continuous(limits = c(0, 1.1))+
  NULL

```

# multiple draws from uniform

In the above example with took a single draw for the trait values from the uniform distribution. In the code below, we repeat this 1000 times to get a distribution of BEF results from multiple draws

### fitness values
```{r}
Traits <- matrix(
  nrow = length(env),
  ncol = length(spec),
  dimnames = list(env, spec))

# take N random draws

set.seed(159)

N <- 1000

Trait_vals <- runif( n = prod( dim( Traits)) * N, 0, 1)

Trait_list <- 
lapply(1:N, function(x) {Traits[1:nrow(Traits), 1:ncol(Traits)] <- Trait_vals[c(((x-1)*25+1) : (x*25))]
return(Traits)})

```


###plot values
```{r}

plot_values_df_list <- 
pblapply(Trait_list, function(TM) {

plot_values <- sapply(env, #for each environment
                      function(x) {
                        lapply(spec_comb, #for each richness level
                               function(y) {
                                 apply(y, 2, function(z){ #for each species combination
                                         mean(sum(TM[x,z]^2) / abs(sum(TM[x,z])))
                                   }
                                 )
                                 }
                        )
                        }
                      )


plot_values_df <- 
  tibble(
    richness = rep( rep(1:specnum, unlist(lapply(spec_comb, ncol))), envnum),
    spec_comb = rep(lapply(spec_comb, 
                           function(x) apply(x, 2, function(y) 
                             paste(y, collapse = " "))) %>% unlist(),
                    envnum),
    environment = rep(env, each = sum(choose(specnum,1:specnum))),
    functioning = unlist(plot_values)
    )

return(plot_values_df)
})
```


###BEF metrics
```{r}

env_metrics_list <- 
  pblapply(plot_values_df_list, function(PV) {
    
    plot_values_wide <-
      PV %>% 
      spread(environment, functioning)

#calculate average functioning for each env combination at each scale and store results in dataframe
env_values <- 
lapply(env_comb, function(x) { #for all scales (1:5)
  apply(x, 2, function(y) { #for all environmental combinations
   
     df <- data.frame( #average functioning for this environment combination
       env_comb = paste0(y, collapse = " "),
       envnum = length(y),
       richness = plot_values_wide$richness,
       functioning = rowMeans(plot_values_wide[,y]),
       stringsAsFactors = FALSE
                     )
  })
  })

env_values <- bind_rows(unlist(env_values, recursive = FALSE))

# calculate BEF metrics for each environment combination
env_metrics <- 
  env_values %>% 
  group_by(env_comb, envnum) %>%
  nest() %>% 
  mutate(transgressive = map_dbl(data, function(x) {
    x[x$richness == specnum,]$functioning / 
      max(x[x$richness == 1,]$functioning)})) 

transgressive <- 
env_metrics %>% 
  select(-data) %>% 
  gather(metric, value, -envnum, -env_comb) %>%
  filter(metric == "transgressive") %>% 
  lm(value ~ envnum, data = .)

beta_slopes <- 
env_values %>% 
  group_by(envnum, env_comb) %>%
  nest() %>% 
  mutate(power_fit = map(data, ~nls(functioning ~ a*richness^b, data = ., start = list(a = 1, b = 1)))) %>%
  mutate(beta = map_dbl(power_fit, function(x) coef(x)[2])) %>%
  lm(beta~envnum, data = .) #%>% 


return(tibble(transgressive = list(transgressive),
              beta_slopes = list(beta_slopes)
              ))
}
)


```

```{r}
Results <- 
 bind_rows(env_metrics_list) %>% 
  gather(metric, value) %>% 
  mutate(slope = map_dbl(map(.$value, coef),"envnum")) %>% 
  mutate(Intercept = map_dbl(map(.$value, coef),"(Intercept)")) %>% 
  select(-value)
```

##Fig 2 - slopes and histograms 

colour
```{r}
library(RColorBrewer)
col_2 <- viridis(6)[3]
```

###Fig 2a
```{r}
env_pred = seq(1,5,0.1)

newdata = tibble(rep = rep(1:nrow(Results), each = length(env_pred)),
                 env_pred = rep(env_pred, nrow(Results)))

Results_pred <- 
  Results %>% 
  mutate(rep = 1:n()) %>% 
  left_join(newdata) %>% 
  mutate(predict = slope*env_pred+Intercept)
 
med_slope <- 
Results %>% 
  mutate(rep = 1:n()) %>% 
  filter(metric == "beta_slopes") %>% 
  slice(1:999) %>% #has to be odd number to find exact media
  mutate(slope_q = 
  case_when(abs(slope - median(slope)) == min(abs(slope - median(slope))) ~ "median",
            abs(slope - quantile(slope, 0.025)) == min( abs(slope - quantile(slope, 0.025))) ~ "low_q",
            abs(slope - quantile(slope, 0.975)) == min( abs(slope - quantile(slope, 0.975))) ~ "high_q")) %>% 
  filter(!is.na(slope_q))

Results_beta <- 
Results_pred %>% 
  filter(metric == "beta_slopes")  

Fig_2_a <- 
  ggplot(Results_beta, aes(x = env_pred, y = predict, group = rep))+
  geom_line(size = 0.5, alpha = 0.1, colour = col_2)+
  geom_line(data = Results_beta %>% filter(rep == pull(filter(med_slope,
                                                   slope_q == "median"), rep)), 
            colour = "darkred", size = 1)+
  theme_classic()+
  labs(y = "beta slope", x = "scale")+
  theme(legend.position = "none",
        axis.ticks.length=unit(-0.25, "cm"),
        axis.text.x = element_text(margin=unit(c(0.5,0.5,0.5,0.5), "cm")),
        axis.text.y = element_text(margin=unit(c(0.5,0.5,0.5,0.5), "cm")) )+
  scale_y_continuous(limits = c(-0.2, NA))


  
save("Fig_2_a", file = here("figures", "Fig_2_a.RData"))

Fig_2_a

```

###Fig 2a - inset
```{r}

quantile_res <- 
Results %>% 
  filter(metric == "beta_slopes") %>% 
  pull(slope) %>% 
  quantile(probs = c(0.025, 0.5, 0.975))

Fig_2_a_i <- 
Results %>% 
  filter(metric == "beta_slopes") %>% 
  ggplot(aes(x = slope))+
  geom_histogram(binwidth = 0.001, fill = col_2)+
  geom_vline(xintercept = 0, linetype = "dashed")+
  annotate(geom = "point", x = quantile_res[2], y = 10)+
  annotate(geom = "errorbarh", xmax = quantile_res[1], 
           xmin = quantile_res[3], y = 10)+
  #geom_point(data = . %>% summarise(beta_slopes = median(beta_slopes)), aes(x = beta_slopes, y = 50))+
  theme_bw()+
  theme(legend.position = "none")+
  labs(y = paste("n out of", N, "draws", sep = " "),
       title = "change in production ~ richness relationship with scale",
       x = "")
  


save("Fig_2_a_i", file = here("figures", "Fig_2_a_i.RData"))

Fig_2_a_i

```

###Fig 2b
```{r}
med_slope <- 
Results %>% 
  mutate(rep = 1:n()) %>% 
  filter(metric == "transgressive")  %>% 
  slice(1:999) %>% #has to be odd number to find exact media
  mutate(slope_q = 
  case_when(abs(slope - median(slope)) == min(abs(slope - median(slope))) ~ "median",
            abs(slope - quantile(slope, 0.025)) == min( abs(slope - quantile(slope, 0.025))) ~ "low_q",
            abs(slope - quantile(slope, 0.975)) == min( abs(slope - quantile(slope, 0.975))) ~ "high_q")) %>% 
  filter(!is.na(slope_q))


Results_trans <- 
Results_pred %>% 
  filter(metric == "transgressive")  

Fig_2_b <- 
ggplot(Results_trans, aes(x = env_pred, y = predict, group = rep))+
  geom_line(size = 0.5, alpha = 0.1, colour = col_2)+
  geom_line(data = Results_trans %>% filter(rep == pull(filter(med_slope, slope_q == "median"), rep)), colour = "darkred", size = 1)+
  theme_classic()+
  labs(y = "transgressive overyielding", x = "scale")+
  theme(legend.position = "none",
        axis.ticks.length=unit(-0.25, "cm"),
        axis.text.x = element_text(margin=unit(c(0.5,0.5,0.5,0.5), "cm")),
        axis.text.y = element_text(margin=unit(c(0.5,0.5,0.5,0.5), "cm")) )+
  scale_y_continuous(limits = c(-0.2, NA))


save("Fig_2_b", file = here("figures", "Fig_2_b.RData"))


Fig_2_b  
```

##Fig 2b inset
```{r}

quantile_res <- 
Results %>% 
  filter(metric == "transgressive") %>% 
  pull(slope) %>% 
  quantile(probs = c(0.025, 0.5, 0.975))

Results_trans <- 
  Results %>% 
  filter(metric == "transgressive") 

Fig_2_b_i <- 
ggplot(Results_trans, aes(x = slope))+
  geom_histogram(binwidth = 0.006, fill = col_2)+
  geom_vline(xintercept = 0, linetype = "dashed")+
  annotate(geom = "point", x = quantile_res[2], y = 6)+
  annotate(geom = "errorbarh", 
           xmax = quantile_res[1], 
           xmin = quantile_res[3],
           y = 6)+
  theme_bw()+
  theme(legend.position = "none")+
  labs(y = paste("n out of", N, "draws", sep = " "),
       title = "change in transgressive overyielding with scale",
       x = "transgressive overyileding ~ scale")+
  scale_x_continuous(limits = c(-0.07, 0.157))


  save("Fig_2_b_i", file = here("figures", "Fig_2_b_i.RData"))
 
Fig_2_b_i
```


##Fig S7 - species specialisation

This code investigates the relationship between transgressive overyielding and the species specialization index.

From the article: 

>[Species specialisation is] calculated as the number of species in monoculture that maximise ecosystem function in at least one habitat divided by the number of habitats and scaled between 0 and 1. For example, if two different species maximise function in two different environments, the index is one. If one species maximises ecosystem function in both environments, the index is zero.


```{r}
Species_specialisation <- 
  lapply(Trait_list, function(x) {
    apply(x, 1, function(y){
      which(y == max(y))
    })
  }) %>% 
  lapply(., function(x) {
    ssp <- length(unique(x))/length(x)
    (ssp-(1/envnum))/(1-(1/envnum))
}) %>% unlist()

sppT <-   
tibble(spec_spec = Species_specialisation, 
       transgressive = filter(Results, metric == "transgressive")$slope)

ggplot(sppT, aes(x = spec_spec, y = transgressive))+
  geom_beeswarm(shape = 21, fill = NA, colour = "#2ca25f")+
  geom_smooth(se = FALSE, method = "lm", colour = "black", size = 0.5)+
  theme_bw()+
  labs(x = "species specialisation index", y = expression(scale - OY[trans]-slope))

ggsave(filename = here("figures", "Fig_S7.pdf"), width = 6, height = 4)


```


### statistic

```{r}

sppT %>% 
  lm(transgressive~spec_spec, data = .) %>% 
  tidy()

```

